---
title: "Are NFL Injuries Getting Better?"
author: "Kristina Lau"
date: "2023-11-20"
output: html_document
theme: modern
highlight: tango
df_print: paged
fig_height: 8
fig_width: 20
number_sections: no
toc: yes
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Introduction

Throughout the NFL's lifespan, there have been many player injuries. Due to concerns over their player's health, the NFL has strived to help lessen the amount of player injuries. One of the ways they have tried to help player safety is by changing their rules. Since 2002 alone, the NFL has made 50 changes to their rules in order to eliminate potentially dangerous tactics and reduce the risk of injuries. Despite all of these rule changes, players continue to suffer serious injuries.

This analysis will go over whether there has been an increase or decrease in player injuries and if there is a solution to decreasing the amount of injuries overall. 

# Packages I used
```{r, echo = TRUE, message=FALSE, warning=FALSE}
# libraries
#helps with data import, tidying, manipulation and data visualization
library(tidyverse)
#to help with dates and times
library(lubridate)
#helps with formatting numbers and labels in a visually appealing way
library(scales)
#helps create complex plots from data and data frames
library(ggplot2)
#allows users to create ridgeline plots in ggplot2
library(ggridges)
```


```{r, message=FALSE, warning=FALSE}
# set up plotting theme
theme_jason <- function(legend_pos="top", base_size=12, font=NA){
  
  # come up with some default text details
  txt <- element_text(size = base_size+3, colour = "black", face = "plain")
  bold_txt <- element_text(size = base_size+3, colour = "black", face = "bold")
  
  # use the theme_minimal() theme as a baseline
  theme_minimal(base_size = base_size, base_family = font)+
    theme(text = txt,
          # axis title and text
          axis.title.x = element_text(size = 15, hjust = 1),
          axis.title.y = element_text(size = 15, hjust = 1),
          # grid lines on plot
          panel.grid.major = element_line(linetype = 2),
          panel.grid.minor = element_line(linetype = 2),
          # title and subtitle text
          plot.title = element_text(size = 18, colour = "grey25", face = "bold"),
          plot.subtitle = element_text(size = 16, colour = "grey44"),
          
          ###### clean up!
          legend.key = element_blank(),
          # the strip.* arguments are for faceted plots
          strip.background = element_blank(),
          strip.text = element_text(face = "bold", size = 13, colour = "grey35")) +
    
    #----- AXIS -----#
    theme(
      #### remove Tick marks
      axis.ticks=element_blank(),
      
      ### legend depends on argument in function and no title
      legend.position = legend_pos,
      legend.title = element_blank(),
      legend.background = element_rect(fill = NULL, size = 0.5,linetype = 2)
      
    )
}


plot_cols <- c("#498972", "#3E8193", "#BC6E2E", "#A09D3C", "#E06E77", "#7589BC", "#A57BAF", "#4D4D4D")
```



## The Data

The two data files I am using are from the website Kaggle. They were provided for an official NFL Data Analysis competition and covers information for:  
  
  * Injury Record: The injury record file in .csv format contains information on 105 lower-limb injuries that occurred during regular season games over the two seasons. Injuries can be linked to specific records in a player history using the PlayerKey, GameID, and PlayKey fields.
* Play List: – The play list file contains the details for the 267,005 player-plays that make up the dataset. Each play is indexed by PlayerKey, GameID, and PlayKey fields. Details about the game and play include the player’s assigned roster position, stadium type, field type, weather, play type, position for the play, and position group.


```{r}
injury_record <- read.csv("C:/Users/kblna/OneDrive/Documents/A UHD Semesters/Fall 2023/Data Science I/git/final-project-fall-2023-kristina-s-team/InjuryRecord - InjuryRecord.csv")
head(injury_record)
play_list <- read.csv("C:/Users/kblna/OneDrive/Documents/A UHD Semesters/Fall 2023/Data Science I/git/final-project-fall-2023-kristina-s-team/PlayList - PlayList.csv")
head(play_list)
```

# Key Terms
### Injury Data File
* PlayerKey - Uniquely identifies a player with a five-digit numerical key
* GameID - Uniquely identifies a player's games(not strictly in temporal order)
* PlayKey - Identifies a player's plays within a game(in sequential order)
* BodyPart - Identifies the injured body part(Knee, Ankle, Foot, etc.)
* Surface - Identifies the playing surface at time of injury(Natural or Synthetic)
* DM_M1 - One-Hot Encoding indicating 1 or more days missed due to injury
* DM_M7 - One-Hot Encoding indicating 7 or more days missed due to injury
* DM_M28 - One-Hot Encoding indicating 28 or more days missed due to injury
* DM_M42 - One-Hot Encoding indicating 42 or more days missed due to injury

### Play List
* PlayerKey - Uniquely identifies a player with a five-digit numerical key
* GameID - Uniquely identifies a player's games(not strictly in temporal order)
* PlayKey - Identifies a player's plays within a game(in sequential order)
* RosterPosition - provides the player's roster position
* PlayerDay - an integer sequence that reflects the timeline of a player's participation in games; use this field to sequence player participation
* PlayerGame - uniquely indentifies a player's games; matches the last integer of the GameID(not strictly in temporal order of game occurrence)
* StadiumType - a free text description of the type of stadium(open, closed dome, etc.)
* FieldType - a categorical description of the field type(Natural or Synthetic)
* Temperature - on-field temperature at the start of the game(not always available - for closed dome/indoor stadiums this field may not be relevant as the temperature and weather are controlled)
* Weather - a free text descroption of the weather at the stadium(for some games this isn't always available)
* PlayType - categorical description of play type(pass, run, kickoff, etc.)
* Position - a categorical variable denoting the player's position for the play(RB, QB, LB, etc.)
* PositionGroup - a categorical variable denoting the player's position group for the position held during the play
***

### Stadium Type

There are `r length(unique(play_list$StadiumType))` different stadium types in the `play_list` dataset.

When we inspect these levels, we can see that there are minor differences in the description of these values, in addition to spelling errors.

```{r, warning=FALSE, message=FALSE}
play_list %>% 
  count(StadiumType) %>% 
  rename(Count = n) %>% 
  mutate(Count = comma(Count)) %>% 
  kableExtra::kable(format = "html", escape = F) %>%
  kableExtra::kable_styling("striped", full_width = F) %>% 
  kableExtra::scroll_box(height = "500px") %>%
  kableExtra::kable_styling(fixed_thead = T)
```


We'll clean these up to five distinct types, and also an additional one for *unknown*:
  
* `outdoor`
* `indoor_closed`
* `indoor_open`
* `dome_closed`
* `dome_open`

```{r, warning=FALSE, message=FALSE}
outdoor <- c('Outdoor', 'Outdoors', 'Cloudy', 'Heinz Field', 
             'Outdor', 'Ourdoor', 'Outside', 'Outddors', 
             'Outdoor Retr Roof-Open', 'Oudoor', 'Bowl')

indoor_closed <- c('Indoors', 'Indoor', 'Indoor, Roof Closed', 'Indoor', 'Roof Closed',
                   'Retractable Roof', 'Retr. Roof-Closed', 'Retr. Roof - Closed', 'Retr. Roof Closed')

indoor_open <- c('Indoor, Open Roof', 'Open', 'Retr. Roof-Open', 'Retr. Roof - Open')

dome_closed <- c('Dome', 'Domed, closed', 'Closed Dome', 'Domed', 'Dome, closed')

dome_open <- c('Domed, Open', 'Domed, open')

convert_stadiums <- function(x) {
  if(x %in% outdoor) {
    "outdoor"
  } else if(x %in% indoor_closed) {
    "indoor closed"
  } else if(x %in% indoor_open) {
    "indoor open"
  } else if(x %in% dome_closed) {
    "dome_closed"
  } else if( x %in% dome_open) {
    "dome_open"
  } else {
    "unknown"
  }
  
}    

play_list <- play_list %>% 
  mutate(StadiumType = mapply(convert_stadiums, StadiumType))
```

### Weather

Similarly the weather.

In the raw data provided, there are `r length(unique(play_list$Weather))` different weather types categorized: 

```{r, warning=FALSE, message=FALSE}
play_list %>% 
  count(Weather) %>% 
  rename(Count = n) %>% 
  mutate(Count = comma(Count)) %>% 
  kableExtra::kable(format = "html", escape = F) %>%
  kableExtra::kable_styling("striped", full_width = F) %>% 
  kableExtra::scroll_box(height = "500px") %>%
  kableExtra::kable_styling(fixed_thead = T)
```


I will be condensing these in to five groups, in addition to "Unknown" for those types where the weather variable is missing:
  
* `rain`
* `overcast`
* `clear`
* `snow`
* `none`


```{r, warning=FALSE, message=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# ----- condense GameWeather variable: -----#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
rain <- c('30% Chance of Rain', 'Rainy', 'Rain Chance 40%', 'Showers', 'Cloudy, 50% change of rain', 'Rain likely, temps in low 40s.',
          'Cloudy with periods of rain, thunder possible. Winds shifting to WNW, 10-20 mph.',
          'Scattered Showers', 'Cloudy, Rain', 'Rain shower', 'Light Rain', 'Rain')

overcast <- c('Party Cloudy', 'Cloudy, chance of rain',
              'Coudy', 
              'Cloudy and cold', 'Cloudy, fog started developing in 2nd quarter',
              'Partly Clouidy', 'Mostly Coudy', 'Cloudy and Cool',
              'cloudy', 'Partly cloudy', 'Overcast', 'Hazy', 'Mostly cloudy', 'Mostly Cloudy',
              'Partly Cloudy', 'Cloudy')

clear <- c('Partly clear', 'Sunny and clear', 'Sun & clouds', 'Clear and Sunny',
           'Sunny and cold', 'Sunny Skies', 'Clear and Cool', 'Clear and sunny',
           'Sunny, highs to upper 80s', 'Mostly Sunny Skies', 'Cold',
           'Clear and warm', 'Sunny and warm', 'Clear and cold', 'Mostly sunny',
           'T: 51; H: 55; W: NW 10 mph', 'Clear Skies', 'Clear skies', 'Partly sunny',
           'Fair', 'Partly Sunny', 'Mostly Sunny', 'Clear', 'Sunny')

snow <- c('Cloudy, light snow accumulating 1-3"', 'Heavy lake effect snow', 'Snow')

none <- c('N/A Indoor', 'Indoors', 'Indoor', 'N/A (Indoors)', 'Controlled Climate')

convert_weather <- function(x) {
  if(x %in% rain) {
    "rain"
  } else if(x %in% overcast) {
    "overcast"
  } else if(x %in% clear) {
    "clear"
  } else if(x %in% snow) {
    "snow"
  } else if(x %in% none) {
    "indoors"
  } else {
    "unknown"
  }
  
}    

play_list <- play_list %>% 
  mutate(Weather = mapply(convert_weather, Weather))

```

***

## Injury Record Data

With almost 86% of the recorded injuries in our dataset, the knees and ankles (46% and 40% respectively) are by far the most frequently injured body part, while the heel is the least frequent, with only one recorded injury.

```{r, message=FALSE, warning=FALSE}
injury_record %>% 
  count(BodyPart) %>% 
  ggplot(aes(x= reorder(BodyPart,n), y= n)) +
  geom_col(fill = plot_cols[1], colour = plot_cols[1], alpha = 0.6) +
  labs(x= "Body Part", y= "Num Injuries") +
  ggtitle("Injuries That Occured On The Lower Limbs", subtitle = "86% of the injury records made up by knees and ankles") +
  coord_flip() +
  theme_jason() +
  theme_minimal()
```


Of the provided data, 54% (57 occurrences) of injuries occurred on Synthetic turf. This is interesting, and suggests that lower-limb injuries may have some relationship with the playing surface, as 58% of the games in the provided dataset are played on natural playing surfaces. Which means that fields with synthetic turf tend to cause more injuries than fields with natural grass. 
  
```{r, message=FALSE, warning=FALSE}
p1 <- play_list %>% 
  distinct(GameID, FieldType) %>% 
  count(FieldType) %>% 
  ggplot(aes(x= FieldType, y= n)) +
  geom_col(fill = plot_cols[3], colour = plot_cols[3], alpha = 0.6) +
  labs(x= "Surface", y= "Games Played") +
  ggtitle("More Games Were Played On Natural Turf", subtitle = "58% of games in our data were played on Natural Turf") +
  scale_y_continuous(labels = comma) +
  coord_flip() +
  theme_jason() +
  theme_minimal()

p2 <- injury_record %>% 
  count(Surface) %>% 
  ggplot(aes(x= Surface, y= n)) +
  geom_col(fill = plot_cols[1], colour = plot_cols[1], alpha = 0.6) +
  labs(x= "Surface", y= "Num Injuries") +
  ggtitle("Slight Difference If By Absolute Terms", subtitle = "Synthetic surfaces have slightly more injuries in absolute terms") +
  coord_flip() +
  theme_jason() +
  theme_minimal()


gridExtra::grid.arrange(p1, p2)
```


This tells us that of the data we've been given to analyse, injuries do occur with greater frequency on synthetic playing surfaces.

Over the whole dataset of over 5,700 games, injuries tended to occur in 1.8% of those games. When we split that between natural and synthetic, we can see that players had a 2.33% chance of becoming injured on synthetic turf vs natural grass where they have a 1.45% chance. 

```{r, message = FALSE, warning=FALSE}
# create a variable to indicate whether an injury occurred 
play_list <- play_list %>% 
  mutate(IsInjured = GameID %in% injury_record$GameID)


# play_list %>% 
#   distinct(GameID, FieldType, IsInjured) %>% 
#   count(IsInjured) %>% 
#   mutate(prop = n/sum(n))

play_list %>% 
  distinct(GameID, FieldType, IsInjured) %>% 
  group_by(FieldType, IsInjured) %>% 
  summarise(n = n()) %>% 
  mutate(perc_injured = n / sum(n)) %>% 
  filter(IsInjured == TRUE) %>% 
  ggplot(aes(x= FieldType, y= perc_injured)) +
  geom_col(alpha = 0.6, fill = plot_cols[1], colour = plot_cols[1]) +
  geom_text(aes(label = percent(perc_injured)), vjust=1, size = 7, colour = plot_cols[8]) +
  geom_hline(yintercept = 0.0182, linetype = 2, colour = plot_cols[3]) +
  annotate(geom = "text", x=0.8, y= 0.0192, label = "Injuries are likely to occur\non 1.8% of all plays", colour = plot_cols[3], size = 5) +
  ggtitle("Natural Grass vs Synthetic Turf", subtitle = "Of over 5,700 plays analysed, injuries occured more frequently on Synthetic") +
  theme_jason() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
  theme_minimal()

```


Of the 42 ankle injuries, 60% occurred on synthetic turf, while the 48 recorded knee injuries occurred with identical frequency on both playing surfaces.

If we look further down the list, feet and toes are direct opposites; 71% of foot injuries occurred on natural playing surfaces, while 86% of toe injuries occurred on Synthetic surfaces. Care should be taken with these two lower-limb parts, as both only have seven recorded injuries events each.

```{r, message=FALSE, warning=FALSE}
injury_record %>% 
  ggplot(aes(x=BodyPart, fill = Surface, colour = Surface)) +
  geom_bar(stat = "count", position = "fill", alpha = 0.6) +
  scale_fill_manual(values = plot_cols[c(2,3)]) +
  scale_colour_manual(values = plot_cols[c(2,3)]) +
  scale_y_continuous(labels = percent) +
  ggtitle("Surface and Body Type Relationship", subtitle = "60% of ankle injuries occur on Synthetic, while\nknee injuries occur with identical frequency on both surfaces") +
  theme_jason() +
  theme(axis.title.y = element_blank()) +
  theme_minimal()
```


```{r, message=FALSE, warning=FALSE}
# create a variable to indicate the severity of the injury (time missed)
injury_record <- injury_record %>% 
  mutate(severity = ifelse(DM_M42 == 1, "42", 
                           ifelse(DM_M42 == 0 & DM_M28 == 1, "28",
                                  ifelse(DM_M42 == 0 & DM_M28 == 0 & DM_M7 == 1, "7", "1"))))
```


Of the seven recorded foot injuries, no player was able to return to the field in less than 28 days.

When focusing on the ankle and knees, the rate at which players returned within the first 28 days was almost identical, with the only difference being in ankle injuries having a slightly higher occurrence of returning within the first 7 days than knees.


```{r, message=FALSE, warning=FALSE}
injury_record %>% 
  # count(Surface, DM_M42) %>% 
  ggplot(aes(x=BodyPart, fill = factor(severity, levels = c("42", "28", "7", "1")), colour = factor(severity, levels = c("42", "28", "7", "1")))) +
  geom_bar(stat = "count", position = "fill", alpha = 0.6) +
  scale_fill_manual(values = plot_cols, name = "Severity") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  scale_y_continuous(labels = percent) +
  ggtitle("Some Differences Between Body Part and Severity", subtitle = "Foot injuries generally have longer recovery times") +
  theme_jason() +
  theme(axis.title.y = element_blank()) +
  theme_minimal()
```


We can see that the time missed due to injury is fairly consistent between the two playing surfaces, so while injuries appear to occur more frequently on synthetic, there doesn't appear to be a pronounced relationship between these two variables.

```{r, message=FALSE, warning=FALSE}
injury_record %>% 
  ggplot(aes(x = Surface, fill = factor(severity, levels = c("42", "28", "7", "1")), colour = factor(severity, levels = c("42", "28", "7", "1")))) +
  geom_bar(stat = "count", position = "dodge", alpha = 0.6) +  # Remove position = "fill"
  scale_fill_manual(values = plot_cols, name = "Severity") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  scale_y_continuous(labels = percent) +
  ggtitle("There Are Minimal Differences in Severity Based On Surface", subtitle = "The time missed because of injury doesn't really differ\nbetween playing surfaces") +
  theme_jason() +
  theme(axis.title.y = element_blank()) +
  theme_minimal()
```


## Injury Record and Play List Data

```{r, message=FALSE, warning=FALSE}
# because there are some injuries without a playID, I will derive the player's position from the play_list df,
# then join on to injury_record
a <- play_list %>% filter(!is.na(Position), Position != "Missing Data") %>% distinct(PlayerKey, Position)

# join to data
injury_record <- injury_record %>% left_join(a, by = "PlayerKey")

# remove the duplicate occurrence of where players have multiple positions
injury_record <- injury_record %>% 
  # mutate(Position = ifelse(is.na(Position), CleanPosition, Position)) %>% 
  distinct(PlayKey, GameID, BodyPart, .keep_all = T) 


# join variables that can be joined at the game level:
injury_record <- injury_record %>% 
  left_join(play_list %>% select(GameID, StadiumType, FieldType, Temperature, Weather), by = "GameID") %>% 
  distinct(.keep_all = T)


# join per play variables
injury_record <- injury_record %>% 
  left_join(play_list %>% select(PlayKey, PlayerDay, PlayerGame, PlayType, PlayerGamePlay), by = "PlayKey")

# fill in missing game number variable where there was no playID in the injury report
injury_record <- injury_record %>% 
  mutate(PlayerGame = ifelse(is.na(PlayerGame), as.numeric(str_extract(GameID, "[^-]*$")), PlayerGame))
```

Wide Receivers, Outside Linebackers and Cornerbacks have the highest number of lower limb injuries with Wide Receivers having the most amount of injuries.

```{r, message=FALSE, warning=FALSE, fig.height=9}
p1 <- injury_record %>% 
  count(Position) %>% 
  ggplot(aes(x= reorder(Position,n), y= n)) +
  geom_col(fill = plot_cols[1], colour = plot_cols[1], alpha = 0.6) +
  ggtitle("Position Injuries", subtitle = "Wide Receivers are the most frequently injured") +
  labs(x= "Position on Play", y= "Injury Count") +
  coord_flip() +
  theme_jason() +
  theme_minimal()
  
p2 <- injury_record %>% 
  count(Position, BodyPart) %>% 
  ggplot(aes(x= reorder(Position,n), y= n, fill = BodyPart, colour = BodyPart)) +
  geom_bar(stat = "identity", position = "fill", alpha = 0.6) +
  scale_fill_manual(values = plot_cols, name = "Body Part") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  scale_y_continuous(labels = percent) +
  ggtitle("Where Each Injury Occured Based On Player Position", subtitle = "Injury location seems to differ based on player's position on the play") +
  labs(x= "Position on Play", y= "Injury Proportion") +
  coord_flip() +
  theme_jason() +
  theme_minimal()

gridExtra::grid.arrange(p1, p2)
```


When analyzing the lower-limb body type injured by position, the main thing you'll notice is of course the high amount of ankle and knee injuries. With Middle Linebackers (MLB) and Safeties being the most likely to injure their ankle than other positions.  

The median play number that players get injured on synthetic playing surfaces is 22, while this number is lower on natural surfaces, with injuries occurring on the 16th play.


```{r, message=FALSE, warning=FALSE}
injury_record %>% 
  ggplot(aes(x= FieldType, y= PlayerGamePlay, fill = FieldType, colour = FieldType)) +
  geom_boxplot(alpha = 0.5) +
  scale_fill_manual(values = plot_cols, guide = "none") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  ggtitle("Injuries Tend to Occur Earlier in the Season", subtitle = "28 records were not included as exact play where injury occurred is unknown") +
  labs(x= "Field Type", y= "Game Play") +
  theme_jason() +
  theme_minimal()
```


We can see that the number of injuries decreases as we progress through the games in the two year period analysed. This would be interesting, however when plot the number of players in the `Play List` data set, we can see that they are also steadily decreasing through the games.

```{r, warning=FALSE, message=FALSE}
p1 <- injury_record %>% 
  group_by(PlayerGame) %>% summarise(num_players = n_distinct(PlayerKey)) %>% 
  ggplot(aes(x=PlayerGame, y=num_players)) +
  geom_line(colour = plot_cols[1], size = 1) +
  geom_point(colour = plot_cols[1], size = 3) +
  labs(x= "Game Number", y= "Number Injuries") +
  ggtitle("The Number of Injuries Per Game Decreases", subtitle = "Week 3 has the most injuries (n=9)") +
  theme_jason() +
  theme_minimal()


p2 <- play_list %>% 
  group_by(PlayerGame) %>% summarise(num_players = n_distinct(PlayerKey)) %>% 
  ggplot(aes(x=PlayerGame, y=num_players)) +
  geom_line(colour = plot_cols[3], size = 1) +
  geom_point(colour = plot_cols[3], size = 3) +
  labs(x= "Game Number", y= "Number Players") +
  ggtitle("Number of Players Injured Per Game Also Decreases") +
  theme_jason() +
  theme_minimal()

gridExtra::grid.arrange(p1, p2, ncol = 1)
```

To further touch upon the issue of whether there is a big difference for the number of games played we can calculate the proportion of injuries (`num_injured` / `num_players`).

When we do this, we can see that based on the data provided, it does appear that players may be more likely to get injured over the earlier games, peaking in week 3 at 3.6%.

```{r, warning=FALSE, message=FALSE}
injury_record %>% 
  group_by(PlayerGame) %>% summarise(num_injured = n_distinct(PlayerKey)) %>% ungroup() %>% 
  left_join(play_list %>% group_by(PlayerGame) %>% summarise(num_players = n_distinct(PlayerKey)) %>% ungroup(), by = "PlayerGame") %>% 
  mutate(prop_injured = num_injured / num_players) %>% 
  ggplot(aes(x=PlayerGame, y=prop_injured)) +
  geom_line(colour = plot_cols[2], size = 1) +
  geom_point(colour = plot_cols[2], size = 3) +
  labs(x= "Game Number", y= "Injury Rate") +
  scale_y_continuous(labels = percent) +
  ggtitle("Players Are More Likely To Be Injured By The 8th Game", subtitle = "The injury rate peaks at week 3 at 3.6%") +
  theme_jason() +
  theme_minimal()
```

There doesn't appear to be much of a relationship between the temperature and whether a player gets injured. The median temperature is ever so slightly higher when the player is injured (62.5 to 61)

```{r}
play_list %>% 
  filter(StadiumType %in% c("dome_open", "indoor open", "outdoor")) %>% 
  filter(Temperature != -999) %>% 
  distinct(GameID, Temperature, IsInjured) %>% 
  ggplot(aes(y= Temperature, x = IsInjured, fill = IsInjured, colour = IsInjured)) +
  geom_boxplot(alpha = 0.5) +
  scale_fill_manual(values = plot_cols, guide = "none") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  labs(x= "Player Injured?") +
  ggtitle("Temperature Isn't as Big of a Deal", subtitle = "There is a slightly higher chance of a player being injured in higher temperatures") +
  coord_flip() +
  theme_jason() +
  theme_minimal()
```


Games played in the rain result in more injuries than any other weather type, with 2% of those plays ending in injury. 

Also of note, of the data provided, no injuries occurred in the snow. 

```{r, warning=FALSE, message=FALSE}
play_list %>% 
  filter(Weather != "indoors") %>% 
  group_by(Weather, IsInjured) %>% 
  summarise(n_Plays = n_distinct(PlayKey)) %>% 
  mutate(proportion_plays = round(n_Plays / sum(n_Plays), 2)) %>% 
  filter(IsInjured == TRUE) %>%
  ggplot(aes(x= reorder(Weather, n_Plays), y= proportion_plays, fill = IsInjured, colour = IsInjured)) +
  geom_col(alpha = 0.6) +
  scale_fill_manual(values = plot_cols, guide = "none") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  geom_text(aes(label = percent(proportion_plays)), hjust= 1, size =5, colour = plot_cols[8]) +
  scale_y_continuous(labels = comma, name = "% Plays") +
  ggtitle("Injuries Are More Likely to Occur In the Rain") +
  labs(caption = "*Indoor stadiums excluded") +
  coord_flip() +
  theme_jason() +
  theme(axis.title.y = element_blank(), axis.text.x = element_blank(), plot.caption = element_text(colour = "darkgrey")) + theme_minimal()
  
```


Games played in indoor stadiums with the retractable roof open appear to result in more injuries than other stadium types, with injuries on 3% of all plays.

```{r, warning=FALSE, message=FALSE}
play_list %>% 
  group_by(StadiumType, IsInjured) %>% 
  summarise(n_Plays = n_distinct(PlayKey)) %>% 
  mutate(proportion_plays = round(n_Plays / sum(n_Plays), 2)) %>% 
  filter(IsInjured == TRUE) %>%
  ggplot(aes(x= reorder(StadiumType, n_Plays), y= proportion_plays, fill = IsInjured, colour = IsInjured)) +
  geom_col(alpha = 0.6) +
  scale_fill_manual(values = plot_cols, guide = "none") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  geom_text(aes(label = percent(proportion_plays)), hjust= 1, size =5, colour = plot_cols[8]) +
  scale_y_continuous(labels = comma, name = "% Plays") +
  ggtitle("Roof Problems", subtitle = "Injuries are more likely to occur in indoor stadiums with the roof open") +
  coord_flip() +
  theme_jason() +
  theme(axis.title.y = element_blank(), axis.text.x = element_blank()) +
  theme_minimal()
```


When we break that out into the different field types, we can see that games played on synthetic turf in indoor stadiums are by far the most problematic, with 7% of the plays on this surface resulting in an injury.

```{r}
play_list %>% 
  group_by(StadiumType, FieldType, IsInjured) %>% 
  summarise(n_Plays = n_distinct(PlayKey)) %>% 
  mutate(proportion_plays = round(n_Plays / sum(n_Plays), 2)) %>% 
  filter(IsInjured == TRUE) %>%
  ggplot(aes(x= reorder(StadiumType, n_Plays), y= proportion_plays, fill = FieldType, colour = FieldType)) +
  geom_col(alpha = 0.6) +
  scale_fill_manual(values = plot_cols, guide = "none") +
  scale_colour_manual(values = plot_cols, guide = "none") +
  geom_text(aes(label = percent(proportion_plays)), hjust= -0, size =5, colour = plot_cols[8]) +
  scale_y_continuous(labels = comma, name = "% Plays") +
  ggtitle("Indoor Stadiums With Synthetic Turf Are The Worst", subtitle = "Injuries are far more likely here, with 7% injury rate") +
  coord_flip() +
  theme_jason() +
  theme(axis.title.y = element_blank(), axis.text.x = element_blank()) + 
  facet_wrap(~ FieldType) +
  theme_minimal()
```

# Synthetic Turf
Now that we've realized that there is a higher chance for players to be injured on Synthetic Turf. Let's go into a brief history lesson about synthetic turf so we can understand it better. 

Synthetic turf, also known as artificial turf, is a specially designed surface composed of synthetic fibers that emulate the appearance and texture of natural grass. Modern iterations of synthetic turf predominantly utilize materials such as nylon, polypropylene, and recycled crumb rubber. Players often attest to experiencing enhanced traction and cutting ability when playing on synthetic turf compared to natural grass. However, this also leads to more lower limb injuries that wouldn't happen otherwise. 

The first NFL team to officially use synthetic turf on their home field was the Seattle Seahawks in 2002. This largely impacted the league as there was an increase in team owners deciding to change their fields in favor of synthetic turf. This was largely due to the owners realizing that synthetic turf is cheaper and easier to manage than natural grass. As of 2021, 15 NFL fields out of 30 use synthetic turf.

***

# Summary

In the NFL, injuries will always happen since that is the nature of sports. Many factors can go into an NFL player's injury, but after looking through the provided data players are more likely to be injured indoors on synthetic turf. While the NFL is always trying to improve player safety, changing the fields to natural grass would greatly help decrease player injuries and be another step in improving player safety.   

 