---
title: "Is Synthetic Turf Worse for NFL Injuries?"
author: "Kristina Lau"
date: "2023-11-30"
output: html_document
fig_height: 10
fig_width: 20
number_sections: no
toc: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

## Introduction


  Throughout the NFL's history, it has always been plagued by injuries. However, the NFL is always striving to improve their player's safety and as a result has made vast improvements when it comes to player healthcare, recovery, and medical procedures. Despite these efforts, players continue to be injured and, in recent seasons, injuries appear to be increasing with synthetic turf being one of the major player complaints. Is synthetic turf increasing the amount of injuries players are receiving in a season? 
  
  
# Questions

The questions I will be asking and attempting to answer are:

* Is there a relationship between injuries and the type of field stadiums use?

* Which positions get injured more often?

* What is the cause of these injuries if any?

* Is there a way to lessen the amount of injuries in the NFL?

## Packages I Used
```{r, echo = TRUE, message=FALSE, warning=FALSE}
#helps tidy data
library(tidyr)
#helps to wrangle data
library(dplyr)
#helps with data import, tidying, manipulation and data visualization
library(tidyverse)
#helps create complex graphs and plots
library(ggplot2)
#will help with dates and times
library(scales)
#to help create visually appealing graphs
library(viridis)
#helps create interactive web applications for one of my tables
library(shiny)

```

## Key Terms
### Injury Data File
* PlayerKey - Uniquely identifies a player with a five-digit numerical key
* GameID - Uniquely identifies a player's games(not strictly in temporal order)
* PlayKey - Identifies a player's plays within a game(in sequential order)
* BodyPart - Identifies the injured body part(Knee, Ankle, Foot, etc.)
* Surface - Identifies the playing surface at time of injury(Natural or Synthetic)
* DM_M1 - One-Hot Encoding indicating 1 or more days missed due to injury
* DM_M7 - One-Hot Encoding indicating 7 or more days missed due to injury
* DM_M28 - One-Hot Encoding indicating 28 or more days missed due to injury
* DM_M42 - One-Hot Encoding indicating 42 or more days missed due to injury

### Play List
* PlayerKey - Uniquely identifies a player with a five-digit numerical key
* GameID - Uniquely identifies a player's games(not strictly in temporal order)
* PlayKey - Identifies a player's plays within a game(in sequential order)
* RosterPosition - provides the player's roster position
* PlayerDay - an integer sequence that reflects the timeline of a player's participation in games; use this field to sequence player participation
* PlayerGame - uniquely indentifies a player's games; matches the last integer of the GameID(not strictly in temporal order of game occurrence)
* StadiumType - a free text description of the type of stadium(open, closed dome, etc.)
* FieldType - a categorical description of the field type(Natural or Synthetic)
* Temperature - on-field temperature at the start of the game(not always available - for closed dome/indoor stadiums this field may not be relevant as the temperature and weather are controlled)
* Weather - a free text descroption of the weather at the stadium(for some games this isn't always available)
* PlayType - categorical description of play type(pass, run, kickoff, etc.)
* Position - a categorical variable denoting the player's position for the play(RB, QB, LB, etc.)
* PositionGroup - a categorical variable denoting the player's position group for the position held during the play
***

## The Data

[Here is the Link for the Data](https://www.kaggle.com/competitions/nfl-playing-surface-analytics/data?select=PlayerTrackData.csv)

The data I will be using is from the website Kaggle. They were provided for an official NFL Data competition and provide information regarding player injuries throughout two seasons. 

 * Injury Record: This csv file contains information on 105 lower-limb injuries that occured during regular season games over two seasons. Injuries can be linked to specific records in a player history using the PlayerKey, GameID, and PlayKey fields.
 * Play List: The play list file contains the details for 267, 005 player-plays that make up the dataset. Each play is indexed by PlayKey, GameID, and PlayKey fields. Details about the game and play include the player's assigned roster position, stadium type, field type, weather, play type, position for the play, and position group. 

```{r, echo = TRUE, message=FALSE, warning=FALSE}

injury_record <- read.csv("InjuryRecord - InjuryRecord.csv")
head(injury_record)
play_list <- read.csv("PlayList - PlayList.csv")
head(play_list)

```

# Cleaning up the Data
I will be filtering out the different types of stadiums and grouping them up with each player. 

```{r, echo = TRUE, message=FALSE, warning=FALSE}
injury_record %>% count(PlayerKey)

```



# Where the injuries happened on the player's body

```{r, echo = FALSE, message=FALSE, warning=FALSE}

ggplot(injury_record, aes(x = BodyPart, fill = BodyPart)) +
  geom_bar() +
  geom_text(aes(label = stat(count), group = 1), 
            stat = "count", 
            vjust = -0.5, 
            position = position_stack(), 
            color = "black", 
            size = 3) +
  labs(title = "Injuries By Body Part", x = "Body Part", y = "Total")
```

Here are the different types of injuries in the dataset. There are 105 total injuries with all of them happening to the lower limbs. What is interesting about this data is that the majority of the injuries happened to the knees and ankles. With both of them making up 86% of the total injuries. This suggests that players are either being tackled below their waist a lot or the possibility that something is going on with their  

I plotted the data and put them into a bar graph so we can see all of the different injuries.
For this bar graph I used BodyPart for my x-axis and the total injuries are the y-axis

# Which positions are being injured the most?

```{r, echo = FALSE, message=FALSE, warning=FALSE}
ggplot(play_list %>% group_by(Position) %>%  count(Position)) +
  geom_col(aes(x = reorder(Position, n), y = n), 
               alpha = (1), position = "stack") +
  coord_flip() + 
  labs(title = "Injuries Based on Position", x = "Position", y = "Total")

```



# Stadiums

Here is a list of all of the different stadiums where the injuries happened. 

```{r, echo = FALSE, message=FALSE, warning=FALSE}
play_list %>% count(StadiumType)
  
```

# Which position is being injured on which surface type? 

In this graph there is a

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(play_list, aes( x = FieldType)) +
  geom_bar() +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "none") +
  ggtitle("Injuries based on field type") +
  theme( 
    legend.position = "none",
    panel.spacing = unit(0, "lines"),
    strip.text.x = element_text(size = 7), 
  ) +
  facet_wrap(~Position, scale = "free_y")

```


# Natural vs Synthetic Turf

According to the data from our dataset, injuries appear to happen more often on synthetic turf than natural turf. 

```{r, echo = FALSE, message=FALSE, warning=FALSE}
ggplot(injury_record, aes(x = Surface, fill = Surface)) +
  geom_bar(width = 0.5, color = "black") +
  geom_text(aes(label = stat(count), group = 1), 
            stat = "count", 
            vjust = -0.5, 
            position = position_stack(), 
            color = "black", 
            size = 3) +
  scale_fill_manual(values = c("Natural" = "forestgreen", "Synthetic" = "darkolivegreen")) +
  scale_color_gradient2(low = "white", high = "green") +
  labs(title = "Injuries Based on Surface", x = "Surface", y = "Total")
```

# Synthetic Turf
Now that we've realized that there is a higher chance for players to be injured on Synthetic Turf. Let's go into a brief history lesson about synthetic turf so we can understand it better. 

Synthetic turf, also known as artificial turf, is a specially designed surface composed of synthetic fibers that emulate the appearance and texture of natural grass. Modern iterations of synthetic turf predominantly utilize materials such as nylon, polypropylene, and recycled crumb rubber. Players often attest to experiencing enhanced traction and cutting ability when playing on synthetic turf compared to natural grass. However, this also leads to more lower limb injuries that wouldn't happen otherwise. 

The first NFL team to officially use synthetic turf on their home field was the Seattle Seahawks in 2002. This made a huge impact on the league as there was an increase in team owners deciding to change their fields in favor of synthetic turf. This was largely due to the owners realizing that synthetic turf is cheaper and easier to manage than natural grass. As of 2021, 15 NFL fields out of 30 use synthetic turf.

However, after researching the true costs of synthetic turf I've discovered that the costs to install and maintain synthetic turfs are actually much higher than it would be to maintain natural grass fields. Because synthetic turf has to be completely removed and 

***

# References



# Summary

I was curious if there was a relationship between NFL player injuries and the stadium fields they played on. 
