---
title: "Is Synthetic Turf Causing an Increase in NFL Injuries?"
author: "Kristina Lau"
date: "2023-11-30"
output: html_document
fig_height: 10
fig_width: 20
number_sections: no
toc: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

## Introduction


  Throughout the NFL's history, it has always been plagued by injuries. However, the NFL is always striving to improve their player's safety and as a result has made vast improvements when it comes to player healthcare, recovery, and medical procedures. Despite these efforts, players continue to be injured and, in recent seasons, injuries appear to be increasing with synthetic turf being one of the major player complaints. Is synthetic turf causing an increase in the amount of injuries players are receiving in a season? 
  
---  

# Questions

The questions I will be asking and attempting to answer are:

* Where on the player's body are the injuries happening the most?

* Which positions get injured more often?

* Is there a relationship between injuries and the type of field stadiums use?

* Is there a way to lessen the amount of injuries in the NFL?

---

## Packages I Used
```{r, echo = TRUE, message=FALSE, warning=FALSE}
#helps tidy data
library(tidyr)
#helps to wrangle data
library(dplyr)
#helps with data import, tidying, manipulation and data visualization
library(tidyverse)
#helps create complex graphs and plots
library(ggplot2)
#will help with dates and times
library(scales)
#to help create visually appealing graphs
library(viridis)
#helps create interactive web applications for one of my tables
library(shiny)

```

---

## Key Terms
### Injury Data File
* PlayerKey - Uniquely identifies a player with a five-digit numerical key
* GameID - Uniquely identifies a player's games(not strictly in temporal order)
* PlayKey - Identifies a player's plays within a game(in sequential order)
* BodyPart - Identifies the injured body part(Knee, Ankle, Foot, etc.)
* Surface - Identifies the playing surface at time of injury(Natural or Synthetic)
* DM_M1 - One-Hot Encoding indicating 1 or more days missed due to injury
* DM_M7 - One-Hot Encoding indicating 7 or more days missed due to injury
* DM_M28 - One-Hot Encoding indicating 28 or more days missed due to injury
* DM_M42 - One-Hot Encoding indicating 42 or more days missed due to injury

### Play List
* PlayerKey - Uniquely identifies a player with a five-digit numerical key
* GameID - Uniquely identifies a player's games(not strictly in temporal order)
* PlayKey - Identifies a player's plays within a game(in sequential order)
* RosterPosition - provides the player's roster position
* PlayerDay - an integer sequence that reflects the timeline of a player's participation in games; use this field to sequence player participation
* PlayerGame - uniquely indentifies a player's games; matches the last integer of the GameID(not strictly in temporal order of game occurrence)
* StadiumType - a free text description of the type of stadium(open, closed dome, etc.)
* FieldType - a categorical description of the field type(Natural or Synthetic)
* Temperature - on-field temperature at the start of the game(not always available - for closed dome/indoor stadiums this field may not be relevant as the temperature and weather are controlled)
* Weather - a free text descroption of the weather at the stadium(for some games this isn't always available)
* PlayType - categorical description of play type(pass, run, kickoff, etc.)
* Position - a categorical variable denoting the player's position for the play(RB, QB, LB, etc.)
* PositionGroup - a categorical variable denoting the player's position group for the position held during the play
***

---

## The Data

[Here is the Link for the Data](https://www.kaggle.com/competitions/nfl-playing-surface-analytics/data?select=PlayerTrackData.csv)

The data I will be using is from the website Kaggle. They were provided for an official NFL Data competition and provide information regarding player injuries throughout two seasons. 

 * Injury Record: This csv file contains information on 105 lower-limb injuries that occured during regular season games over two seasons. Injuries can be linked to specific records in a player history using the PlayerKey, GameID, and PlayKey fields.
 * Play List: The play list file contains the details for 267, 005 player-plays that make up the dataset. Each play is indexed by PlayKey, GameID, and PlayKey fields. Details about the game and play include the player's assigned roster position, stadium type, field type, weather, play type, position for the play, and position group. 
 
 
For these datasets, I will be reading them in and assigning them to new variables. I will also be displaying the heads of the datasets. 
```{r, echo = TRUE, message=FALSE, warning=FALSE}

injury_record <- read.csv("InjuryRecord - InjuryRecord.csv")
head(injury_record)
play_list <- read.csv("PlayList - PlayList.csv")
head(play_list)

```
---

# Cleaning up the Data

I manually cleaned up the data in the .csv files by fixing spelling mistakes as there were many typos.

I will be filtering out the different types of stadiums and grouping them up with each player. I first did this by using the count function to show each different stadium type.

```{r, message=FALSE, warning=FALSE}
play_list %>% count(StadiumType)
```

Here we can see that many of the different stadium types are primarily either indoors or outdoors. As such I will be separating them into 2 different vectors called 'indoors' and 'outdoors'. Any stadium that is inputted as open or closed I will separate into either vector accordingly.  

Something of note is that there is a strange category called 'Cloudy' which is more than likely a mistake when inputting the data. There are also missing entries for the stadium type and as such I will have to skip over those since there is no way for me to know which stadium type it is. As for the 'Heinz Field' category this stadium is an outdoor stadium so it will go into 'outdoors'.   

```{r, echo = FALSE, message=FALSE, warning=FALSE}

indoors <- c("Bowl", "Closed Dome", "Dome, closed", "Domed", "Domed, closed", "Indoors, Roof Closed", "Retr. Roof - Closed", "Retr. Roof Closed", "Retr. Roof-Closed", "Retractable Roof", "Indoors")

outdoors <- c("Heinz Field", "Domed, open", "Domed, Open", "Indoors, Open Roof", "Open", "Outdoors", "Outdoors Retr Roof-Open", "Retr. Roof - Open", "Retr. Roof-Open")

```

  While I was creating one of my graphs, I realized that they had separated the Safety position into: Free Safety (FS), Strong Safety (SS), and Safety (S). So I will be mutating them into the same position 'Safety'. 

```{r}
#play_list <- mutate(Safety = FS, SS, S)

```


---

# Where on the player's body are the injuries happening the most?

```{r, echo = FALSE, message=FALSE, warning=FALSE}

ggplot(injury_record, aes(x = BodyPart, fill = BodyPart)) +
  geom_bar() +
  geom_text(aes(label = stat(count), group = 1), 
            stat = "count", 
            vjust = -0.5, 
            position = position_stack(), 
            color = "black", 
            size = 3) +
  labs(title = "Injuries By Body Part", x = "Body Part", y = "Total")
```

I plotted the data and put them into a bar graph so we can see all of the different injuries. For this bar graph I used 'BodyPart' for my x-axis and the total injuries are the y-axis.

Here are the different types of injuries in the dataset. There are 105 total injuries with all of them happening to the lower limbs. What is interesting about this data is that the majority of the injuries happened to the knees and ankles. With both of them making up 86% of the total injuries. The fact that all of the injuries are lower limb injuries suggests that there could be a relationship between player injuries and the field types they play on. But so far we don't have anything substantial to prove anything that connects player injuries to the field types. 

---

# Which positions get injured more often?

```{r, echo = FALSE, message=FALSE, warning=FALSE}
ggplot(play_list %>% group_by(Position) %>%  count(Position)) +
  geom_col(aes(x = reorder(Position, n), y = n), 
               alpha = (1), position = "stack") +
  coord_flip() + 
  labs(title = "Injuries Based on Position", x = "Position", y = "Total")

```

I created a horizontal bar graph to analyze the frequency of injuries across different positions. The y-axis represents 'Position,' while the x-axis corresponds to the total number of injuries. 

Upon examining the graph, it is evident that wide receivers experience the highest frequency of injuries while punters experience the lowest frequency of injuries. What is interesting to note is that wide receivers, outside linebackers and cornerbacks are the top 3 positions that tend to be injured the most. Wide receivers and cornerbacks are usually paired up together with wide receivers trying to gain yards and cornerbacks trying to stop them. This means that there are plenty of opportunities for injuries between the two positions to happen.  

---

# Is there a relationship between injuries and the type of field stadiums use? 

## Natural grass vs Synthetic Turf

  I will be comparing the injuries that happen on natural grass vs synthetic turf. This will allow us to see which type of field the injuries are happening the most on. For this comparison I will be plotting a bar graph and using the 'Surface' category as the x-axis and then using the total amount of injuries as the y-axis.  

```{r, echo = FALSE, message=FALSE, warning=FALSE}
ggplot(injury_record, aes(x = Surface, fill = Surface)) +
  geom_bar(width = 0.5, color = "black") +
  geom_text(aes(label = stat(count), group = 1), 
            stat = "count", 
            vjust = -0.5, 
            position = position_stack(), 
            color = "black", 
            size = 3) +
  scale_fill_manual(values = c("Natural" = "forestgreen", "Synthetic" = "darkolivegreen")) +
  labs(title = "Injuries Based on Surface", x = "Surface", y = "Total")
```

According to the data from our dataset, injuries appear to happen more often on synthetic turf than natural grass. With 57% of the injuries happening on synthetic turf. 

Now we're getting somewhere. This helps to prove that of the 105 injuries in the dataset 

---

# Injuries per Position

  For this graph, I wanted to see which positions were being injured on which field type. I created a facet wrap for each position with the x-axis as the field type and the count as the y-axis.
  
```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(play_list, aes( x = FieldType)) +
  geom_bar() +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "none") +
  ggtitle("Injuries based on field type") +
  theme( 
    legend.position = "none",
    panel.spacing = unit(0, "lines"),
    strip.text.x = element_text(size = 7), 
  ) +
  facet_wrap(~Position, scale = "free_y")

```

Now this is very interesting. This heavily contradicts with the previous graph since the majority of the injuries were happening on synthetic turf. And yet for this graph its saying that there are more injuries happening on natural grass. 

---
# Synthetic Turf

Now that we've realized that there is a higher chance for players to be injured on Synthetic Turf. Let's go into a brief history lesson about synthetic turf so we can understand it better. 

Synthetic turf, also known as artificial turf, is a specially designed surface composed of synthetic fibers that emulate the appearance and texture of natural grass. Modern iterations of synthetic turf predominantly utilize materials such as nylon, polypropylene, and recycled crumb rubber. Players often attest to experiencing enhanced traction and cutting ability when playing on synthetic turf compared to natural grass. However, this also leads to more lower limb injuries that wouldn't happen otherwise. 

The adoption of synthetic turf in the NFL gained prominence when the Seattle Seahawks became the first team to install it on their home field in 2002. This move sparked a trend, leading to more team owners opting for synthetic turf. As of 2021, 15 out of 30 NFL stadiums feature synthetic turf. It's worth noting that the shared stadiums of the New York Giants and Jets, as well as the Los Angeles Chargers and Rams, contribute to the total count.

Despite the rising popularity of synthetic turf, research reveals that the installation and maintenance costs exceed those of natural grass fields. This prompts the question: Why do more NFL stadiums choose synthetic turf over natural grass? Because its easier to maintain and lasts longer than natural grass. 

Owners favoring synthetic turf argue that it proves cost-efficient and enhances player performance. The surface is easier to maintain and boasts a longer lifespan than natural grass. Despite concerns about an increased risk of injuries, owners who believe in the benefits of synthetic turf are likely to continue its use, sometimes against the preferences of players and other stakeholders.

---
# Is there a way to lessen the amount of injuries in the NFL?

Yes, there is a way to lessen the amount of injuries in the NFL and that is for all NFL stadiums to use natural grass.  

---
# Summary

I was curious if synthetic turf was the cause of an increase in NFL injuries. For this analysis I first tried to look at what were the most common types of injuries with knee injuries being the highest. Next, I 
